Linear Variation and an Optimization of Algorithms for Connected Components Labeling in a Binary Image

### 2016-04-14 paper/paper.tex
Fix references

\institute{Moscow Institute of Physics and Technology State University, Dolgoprudny, Russia\\
\email{alekseev@phystech.edu}
\and
Chelyabinsk State University, Chelyabinsk, Russia
\and
\email{alexeev@csu.ru}
\and
\email{artemmac@mail.ru}}

@@

\begin{abstract}
The linear variation is a topological characteristic of the function of two variables.
The problem of linear variation computing can be reduced to the problem of counting connected components in a binary image with eight-connected connectivity.
%The new hybrid method for the problem is presented.
%The method is essentially a modification of a two-scan algorithm for labeling
%that groups the pixels into $2 \times 2$ cells.
This is usually done through connected components labeling, and many approaches for that are known.
This paper presents an efficient way to convert the initial binary image with 8-connectivity into a condensed (non-binary) image with new connectivity that is 2 times smaller in each dimension than the initial image.
Most approaches for connected components labeling are still valid for the new image.
As time and memory consumptions of conventional approaches usually depend linearly on the number of pixels,
running them on the condensed image can be up to 4 times more efficient than running on the initial image.
The method is essentially a modification of some known raster algorithms for labeling that groups pixels into $2 \times 2$ cells.
A performance benchmark of the proposed  method applied to some approaches on noise images is provided.

@@

\section{$2\times2$ condensation}

@@

The condensed image will be of size $N \times M$.
Each pixel of the condensed image will unambiguously represent the configuration
of the corresponding $2\times 2$ cell of pixels of the initial binary image.
So the condensed image will contain pixels of $2^4 = 16$ colors.
We suggest the function $c(x, y)$ denoting the color of a pixel of condensed image
to be the following:
\begin{align*}
  c(x, y) &= 2^0 b(2x, 2y)   + 2^1 b(2x+1, 2y) \\
          &+ {} 2^2 b(2x, 2y+1) + 2^3 b(2x+1,2y+1)
\end{align*}

@@

The idea is that if we naturally encode mask pixels as 4-bit numbers in the
way similar to how we defined $c$, we can express the predicate denoting whether
two nearby cells
are directly connected efficiently with just two bitwise and one logical AND
operations.
As an example, \autoref{fig:connectivity:code} compares possible codes in \C checking
if the labels of the current pixel or cell and the one right on top of it
should be same for classic 8-connectivity and for the new connectivity.
Note that as bitwise operations are usually relatively cheap, we added only a
little complication compared to decreasing the number of pixels to process by the 
factor of 4.

@@

Our benchmarks show that, although SUF2 without condensation is generally more
efficient on noise images than SUF without condensation, the mentioned
consideration make the $2 \times 2$ optimization
of SUF2 less dramatic in means of performance increase factor.
Optimized SUF2 significantly outperforms optimized SUF only on
huge noise images with less than 60\% density, while falling behind on
small and medium-sized noise images.

>>>>

\institute{Moscow Institute of Physics and Technology, Dolgoprudny, Russia\\
\email{alekseev@phystech.edu}
\and
Chelyabinsk State University, Chelyabinsk, Russia
\and
\email{alexeev@csu.ru}
\and
\email{artemmac@mail.ru}}

@@

\begin{abstract}
The linear variation is a topological characteristic of the function of two
variables. The problem of linear variation computing can be reduced to the
problem of counting connected components in a binary image with eight-connected
connectivity.
%The new hybrid method for the problem is presented. The method is essentially
%a modification of a two-scan algorithm for labeling that groups the pixels
%into $2 \times 2$ cells.
This is usually done through connected components labeling, and many approaches
for that are known. This paper presents an efficient way to convert the initial
binary image with 8-connectivity into a condensed (non-binary) image with new
connectivity that is 2 times smaller in each dimension than the initial image.
Most approaches for connected components labeling are still valid for the new
image. As time and memory consumptions of conventional approaches usually
depend linearly on the number of pixels, running them on the condensed image
can be up to 4 times more efficient than running on the initial image. The
method is essentially a modification of some known raster algorithms for
labeling that groups pixels into $2 \times 2$ cells. A performance benchmark of
the proposed  method applied to some approaches on noise images is provided.

@@

\section{$2 \times 2$ condensation}

@@

The condensed image will be of size $N \times M$.
Each pixel of the condensed image will unambiguously represent the configuration
of the corresponding $2 \times 2$ cell of pixels of the initial binary image.
So the condensed image will contain pixels of $2^4 = 16$ colors.
We suggest the function $c(x, y)$ denoting the color of a pixel of condensed
image to be the following:
\begin{align*}
  c(x, y) &= 2^0 b(2x, 2y)   + 2^1 b(2x+1, 2y) \\
          &+ {} 2^2 b(2x, 2y+1) + 2^3 b(2x+1,2y+1)
\end{align*}

@@

The idea is that if we naturally encode mask pixels as 4-bit numbers in the
way similar to how we defined $c$, we can express the predicate denoting whether
two nearby cells
are directly connected efficiently with just two bitwise and one logical AND
operations.
As an example, \autoref{fig:connectivity:code} compares possible codes in \C
checking if the labels of the current pixel or cell and the one right on top of
it should be same for classic $8$-connectivity and for the new connectivity. Note
that as bitwise operations are usually relatively cheap, we added only a little
complication compared to decreasing the number of pixels to process by the
factor of $4$.

@@

Our benchmarks show that, although SUF2 without condensation is generally more
efficient on noise images than SUF without condensation, the mentioned
consideration make the $2 \times 2$ optimization of SUF2 less dramatic in means
of performance increase factor. Optimized SUF2 significantly outperforms
optimized SUF only on huge noise images with less than $60\%$ density, while
falling behind on small and medium-sized noise images.
